Making Sequential Programs Deterministic

### 2019-04-06 paper.tex
Edits in abstract

%We have created a type system for determinism.  
We have created a type system that expresses determinism specifications
in a program.
The key ideas in the type system are type qualifiers for nondeterminism,
order-nondeterminism, and determinism; type well-formedness rules to
restrict the typings for collections; and enhancements to polymorphism that
improve precision when analyzing collection operations. While state-of-the-art
nondeterminism detection tools rely on observing runtime output, our approach
aims to verify determinism thereby providing stronger soundness guarantees.

We implemented our type system for Java.
Our type checker, \theDeterminismChecker, warns if a
program is nondeterministic or verifies that the program is deterministic.
In a case study of a 24,000 line software project, it found
previously-unknown nondeterminism errors in a program that had been heavily
vetted by its developers,
who were greatly concerned about nondeterminism errors.
In an experiment, it found all of the
non-concurrency-related nondeterminism that was found by
state-of-the-art dynamic approaches to detecting flaky tests.

@@

%% Keywords
%% comma separated list
\keywords{\todo{keyword1}, \todo{keyword2}, \todo{keyword3}}  %% \keywords are mandatory in final camera-ready submission

>>>>

%We have created a type system for determinism.  
We have created a type system that expresses determinism specifications
in a program.
The key ideas in the type system are type qualifiers for nondeterminism,
order-nondeterminism, and determinism; type well-formedness rules to
restrict the typings for collections; and enhancements to polymorphism that
improve precision when analyzing collection operations. While state-of-the-art
nondeterminism detection tools rely on observing runtime output, our approach
aims to verify determinism at compile time, thereby providing stronger soundness guarantees.

We implemented our type system for Java.
Our type checker, \theDeterminismChecker, warns if a
program is nondeterministic or verifies that the program is deterministic.
In a case study of a 24,000-line software project, it found
previously-unknown nondeterminism errors in a program that had been heavily
vetted by its developers,
who were greatly concerned about nondeterminism errors.
In an experiment, it found all of the
non-concurrency-related nondeterminism that was found by
state-of-the-art dynamic approaches to detecting flaky tests.

@@

%% Keywords
%% comma separated list
\keywords{nondeterminism, type system, verification, specification, hash table}  %% \keywords are mandatory in final camera-ready submission
